On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-02-05 main.tex
Updated Section 4 - Noether position

% packages 
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage{enumitem}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\bz{\textit{\textbf{0}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\mA{{\bm A}}
\def\fA{{\frak A}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\pb{\textit{\textbf{p}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\pr{\mathbb{P}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\K{\mathbb{K}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textbf{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\kt{\widetilde{k}}
\def\dt{s}
\def\bI{\textbf{I}}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\softO}{{O^{\sim}}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\scrQ{\ensuremath{\mathscr{Q}}}
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

We will also work with change of variables with ``generic entries''.
For this, we will introduce $n^2$ new indeterminates $(\frak
A_{i,j})_{1\le i,j \le n^2}$, we will let $\K$ be the field $\C(\frak
A_{i,j})_{1\le i,j \le n^2})$, and use the corresponding matrix $ \fA$
over $\K$. For $g$ as above, we will then define the polynomial 
$g^\fA:=g(\fA \Xb) \in \K[X_1,\dots,X_n]$.

@@

It will sometimes be useful for us use in replace of the matrix $A$, a
matrix representing indeterminants: let $\A$ be an $n \times n$ matrix
of new indeterminates $(\A_{i,j})_{1 \leq i , j \leq n}.$ Define
$f^{\A} \in \mathbb{Q}(\A_{i,j})[\textbf{X}]$ as $f(\A\textbf{X})$ and
$V^\A=V(f^\A).$ It follows from Proposition 5.1, that for a generic $A
\in \C^{n^2},$ $0$ is a regular value of $\Phi_{A}$.  The proof of
\cite[Theorem B.3]{TWT}, which establishes the existence of $\Delta
\subset \C^{n^2},$ applies \textit{Sard's lemma}, which shows that the
critical values of a polynomial mapping are contained in a
hypersurface. It follows from Sard's lemma that, for the matrix of
indeterminants $\A,$ $0$ is a regular value of $\Phi_{\A}.$ Otherwise
it could not be that $0$ is a regular value of $\Phi_A,$ for a generic
$A\in \C^{n^2}.$ The proof is routine and we omit details. It then
follows that
\[
\jac_{\xb}(\Phi_{\A}) = \jac_{\xb}\left(f^{\A}(\xb),\frac{\pa f^{\A}(\xb)}{\pa X_{i+1}},\hdots,\frac{\pa f^{\A}(\xb)}{\pa X_n}\right)
\]
has full rank for all $\xb \in V(f^{\A})$, and thus $f^{\A}$satisfies
$\textbf{H}_i^{'}(1),\textbf{H}_i^{'}(2)$ and $\textbf{H}_i^{'}(3)$ by
the Jacobian criterion \cite[Theorem 16.19]{ECA}. We will rely on this
fact in Section 8.

@@

Throughout this section,  $i \in \{1,\hdots,n\}$ is
fixed. We prove that there exists a hypersurface $\D_{3,i} \subset
\C^{n^2}$ such that if $\mA$ is not in $\D_{3,i}$, then $\mA$ is
invertible and satisfies $\textbf{H}_i$ and $\textbf{H}_i$.

Since $i$ is fixed, to simplify notation, let $\I^\A$ denote
$\I(\pi_i,V^\A)$ and let $W^\A$ denote $W(\pi_i,V^\A).$

Consider again the matrix of indeterminates $\A$.  Define $f^{\A} \in
\mathbb{Q}(\A)[\textbf{X}]$ as $f(\A\textbf{X})$ and $V^\A=V(f^\A).$
Recall from Section~\ref{xxx} that $f^\A$ satisfies
$\textbf{H}_i^{'}(1)$, that is, \td.

\[
\C[\textbf{X}_{\leq i-1}]\rightarrow\C[\textbf{X}]/\frak{I}(\pi_i,V)
\]
injective and  integral.
%

It has already been shown in \cite{EMP} that $f^A$ satisfies
$\textbf{H}_i^{'}(7)$ for a generic change of variables. The new
contribution is in upper bounding the degree of $\Delta_3.$

\begin{lemma} 
The ring
extension \[\mathbb{Q}(\frak{A})[X_1,\dots,X_{i-1}]\rightarrow\mathbb{Q}(\frak{A})[X_1,\dots,X_n]/\I^{\A}\]
\end{lemma}

@@

Let $P_{j} \in \mathbb{Q}(\frak{A})[\Xb_{\leq i-1}][T]$ be as in the proof of Proposition 7.2. Since $P_{j}(X_j) \in\I^{\A}= \langle f^\A,\frac{\partial f^\A}{\partial X_{i+1}},\hdots,\frac{\partial f^\A}{\partial X_n} \rangle,$

@@

For each $P_j,$ there exists a Zariski open set $\Gamma_j \subset \C^{n^2}$ such that if $A \in \Gamma_j$,
then $A$ cancels none of the denominators of $a_{j}$ or $b_{j,l}$. Define $h_j \in \Q[\A]$ to be the product of the denominators of the coefficients of $a_j$ and $b_{j,l}$. Hence, if $A \in \Q^{n^2}-V(h_j)$ then $A$ cancels none of the denominators of $a_{j}$ or $b_{j,l},$ which are providing the integral dependence relation. Now we define 

@@

Hence, if $A\in \Q^{n^2}-\Xi_i$ then

@@

\mathbb{C}[{\Xb}_{\leq i-1}]\rightarrow\mathbb{C}[\Xb]/\I(\pi_i,V^{A})

@@

is integral. We can obtain bounds for $h_j$ by applying an effective Nullstellensatz theorem. However, the theorem requires $P_{j}$ to be a polynomial. We can obtain a polynomial from $P_{j}$ by clearing denominators. We can multiply $P_{j}$ by the least common multiple of its denominators, which we will denote by $d_j \in \Q[\A_{i,j}]$. Put 

@@

$\deg \pjb\leq (D+1)^{n-i+1}.$ 

@@

&\deg \pjb \leq (D+1)^{n-i+1}.&(2D)^{n-i+1}\textrm{?} 

@@

\right\}\leq D,

@@

\right\} \leq D,

@@

    &\left(\prod_{k=1}^{n-i+1} D\right)\Dt_j 
    + (n-i+1)\left(\Dt_j+1\right)\left(\prod_{k=1}^{n-i}D \right)D\\
    = &\left(D^{n-i+1}\right)\Dt_j+
    (n-i+1)\left(\Dt_j+1\right)\left(D^{n-i+1} \right)\\
    &\leq D^{n}\Dt_j + n\left(\Dt_j+1\right)D^{n}\\
    &=D^n\left(\Dt_j + n\left(\Dt_j+1\right) \right).

@@

    &\leq \left(D^n\left(\Dt_j + n\left(\Dt_j+1\right) \right)\right)\Dt_j\\
    &=D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right).

@@

    &\leq  \sum_{j=i}^n \left( D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) \right) \\
    &=  (n-i+1)\left( D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) \right)\\
    &\leq  n D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) .

@@

    &\leq n^2 D^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right),

>>>>

% packages 
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{mathrsfs}
\usepackage{bm}
\usepackage{enumitem}
% new commands and shortcuts
\def\td{{\bf todo}}
\def\bs{\textit{\textbf{s}}}
\def\bz{\textit{\textbf{0}}}
\def\Lb{\textit{\textbf{L}}}
\def\Xb{\textit{\textbf{X}}}
\def\Lambdab{\bm{\Lambda}}
\def\Thetab{\bm{\Theta}}
\def\thetab{\bm{\vartheta}}
\def\mA{{\bm A}}
\def\fA{{\frak A}}
\def\xb{\textit{\textbf{x}}}
\def\fb{\textit{\textbf{f}}}
\def\ab{\textit{\textbf{a}}}
\def\pb{\textit{\textbf{p}}}
\def\ajb{\overline{a_j}}
\def\bjb{\overline{b_{j,l}}}
\def\pjb{\overline{P_{j}}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\htt}{ht}
\def\C{\mathbb{C}}
\def\pr{\mathbb{P}}
\def\vt{\vartheta}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\K{\mathbb{K}}
\def\rs{\mathscr{R}}
\def\P{\mathscr{P}}
\def\d{\delta}
\def\sing{ \textrm{sing}}
\def\codim{ \textrm{codim}}
\def\jac{ \textbf{jac}}
\def\exp{\textrm{exp}}
\def\grad{\textrm{grad}}
\def\rank{\textrm{rank}}
\def\reg{\textrm{reg}}
\def\jt{\widetilde{J}}
\def\At{\widetilde{A}}
\def\Yt{\widetilde{Y}}
\def\dtt{\widetilde{d}}
\def\kt{\widetilde{k}}
\def\dt{s}
\def\bI{\textbf{I}}
%% \def\dt{\widetilde{d}}
\def\Dt{\widetilde{D}}
\def\rk{\textrm{rank }}
\def\pa{\partial}
\def\D{\Delta}
\def\Z{\frak{Z}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\softO}{{O^{\sim}}}
\def\I{\frak{I}}
\def\Is{\frak{I}^{\star}}
\def\A{\frak{A}}
\def\fp{\frak{P}}
\def\sp{\mathscr{P}}
\def\la{\langle}
\def\ra{\rangle}
% matrices 
\def\scrQ{\ensuremath{\mathscr{Q}}}
\def\bbm{\begin{bmatrix}}
\def\ebm{\end{bmatrix}}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

@@

We will also work with changes of variables with ``generic entries''.
For this, we will introduce $n^2$ new indeterminates $(\frak
A_{i,j})_{1\le i,j \le n^2}$, we will let $\K$ be the field $\C(\frak
A_{i,j})_{1\le i,j \le n^2}$, and use the corresponding matrix $ \fA$
over $\K$. For $g$ as above, we will then define the polynomial 
$g^\fA:=g(\fA \Xb) \in \K[X_1,\dots,X_n]$.

@@

Recall that we work with changes of variables with ``generic entries'' of new indeterminants $(\frak
A_{i,j})_{1\le i,j \le n^2}$, and define the polynomial 
$f^\fA:=f(\fA \Xb)$. Note that $f^\A$ satisfies $\textbf{H}_i(1).$ Indeed, otherwise it could not be that that the Jacobian of $f^A,\frac{\pa f^A}{\pa X_{i+1}},\hdots,\frac{\pa f^A}{\pa X_n}$ has full rank at all $\xb \in V(f^A)$ for a generic $A \in \C^{n^2},$ which would contradict Popositon 3.1. The proof of this statement is routine and we omit details. We will rely on this
fact in Section 8.

@@

Fix $i \in \{1,\hdots,n\}$. Throughout this section, we will avoid writing $i$ where the context is clear. Let $\I^{\A}$ denote $\I(i,V^{\A}).$

We prove that there exists a hypersurface $\D_{3,i} \subset
\C^{n^2}$ such that if $\mA$ is not in $\D_{3,i}$, then $\mA$ is
invertible and satisfies $\textbf{H}_i(1)$ and $\textbf{H}_i(2)$.

Recall that we work with changes of variables with ``generic entries'' of new indeterminants $(\frak
A_{i,j})_{1\le i,j \le n^2}$, and define the polynomial 
$f^\fA:=f(\fA \Xb)$. Recall as well that $f^\A$ satisfies $\textbf{H}_i(1),$ as was stated at the end of Section 3.1. 
\subsection{Sketch of the proof}
Let $f \in \ZZ[X_1,\hdots,X_n]$ have total degree $d$ with $V(f) \subset \C^n$ smooth. Recall $\textbf{H}_i(2):$ \textit{$W(\pi_i,V(f))$ is either empty or in Noether position for
  $\pi_{i-1}$}, which is equivalent to having the extension  
\C[\textbf{X}_{\leq i-1}]\rightarrow\C[\textbf{X}]/\frak{I}(\pi_i,V)
injective and  integral.
\begin{theorem}
There exists a hypersurface $\Delta_3 \subset \C^{n^2}$ of degree at most $\td$ with the property that, if $A \in \C^{n^2} - \Delta_3$ then, for each $i \in\{1,\hdots,n\}, \mA$ is
invertible and satisfies $\textbf{H}_i(1)$ and $\textbf{H}_i(2)$.
\end{theorem}
The ring extension \[\mathbb{Q}(\frak{A})[\textbf{X}_{\leq i-1}]\rightarrow\mathbb{Q}(\frak{A})[\textbf{X}]/\I^{\A}\] is integral.

@@

Let $P_{j} \in \mathbb{Q}(\frak{A})[\Xb_{\leq i-1}][T]$ be as in the proof of Proposition 4.2. Since $P_{j}(X_j) \in\I^{\A}= \langle f^\A,\frac{\partial f^\A}{\partial X_{i+1}},\hdots,\frac{\partial f^\A}{\partial X_n} \rangle,$

@@

For each $P_j,$ there exists a Zariski open set $\Gamma_j \subset \C^{n^2}$ such that if $\mA \in \Gamma_j$,
then $\mA$ cancels none of the denominators of $a_{j}$ or $b_{j,l}$. Define $h_j \in \Q[\A]$ to be the product of the denominators of the coefficients of $a_j$ and $b_{j,l}$. Hence, if $\mA \in \C^{n^2}-V(h_j)$ then $\mA$ cancels none of the denominators of $a_{j}$ or $b_{j,l},$ which are providing the integral dependence relation. Now we define 

@@

Hence, if $\mA\in \Q^{n^2}-\Xi_i$ then

@@

\mathbb{C}[{\Xb}_{\leq i-1}]\rightarrow\mathbb{C}[\Xb]/\I(\pi_i,V^{\mA})

@@

is integral. We can obtain bounds for $h_j$ by applying an effective Nullstellensatz theorem. However, the theorem requires $P_{j}$ to be a polynomial. We can obtain a polynomial from $P_{j}$ by clearing denominators. We can multiply $P_{j}$ by the least common multiple of its denominators, which we will denote by $d_j \in \Q[\A]$. Put 

@@

$\deg \pjb\leq (d+1)^{n-i+1}.$ 

@@

&\deg \pjb \leq (d+1)^{n-i+1}.&(2d)^{n-i+1}\textrm{?} 

@@

\right\}\leq d,

@@

\right\} \leq d,

@@

    &\left(\prod_{k=1}^{n-i+1} d\right)\Dt_j 
    + (n-i+1)\left(\Dt_j+1\right)\left(\prod_{k=1}^{n-i}d \right)d\\
    = &\left(d^{n-i+1}\right)\Dt_j+
    (n-i+1)\left(\Dt_j+1\right)\left(d^{n-i+1} \right)\\
    &\leq d^{n}\Dt_j + n\left(\Dt_j+1\right)d^{n}\\
    &=d^n\left(\Dt_j + n\left(\Dt_j+1\right) \right).

@@

    &\leq \left(d^n\left(\Dt_j + n\left(\Dt_j+1\right) \right)\right)\Dt_j\\
    &=d^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right).

@@

    &\leq  \sum_{j=i}^n \left( d^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) \right) \\
    &=  (n-i+1)\left( d^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) \right)\\
    &\leq  n d^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right) .

@@

    &\leq n^2 d^n\Dt_j\left(\Dt_j + n\left(\Dt_j+1\right) \right),
